---
title: "Available Building Permit Data"
format: html
editor: visual
execute: 
  message: false
  warning: false
---

```{r}
#| label: setup
library(readxl)
library(readr)
library(here)
library(dplyr)
library(sf)
```

## Introduction

The general idea was to see if we can use published building permits to identify local projects of densification to see if respondents in our Ontario 2022 survey who have experienced densification in their neighbourhood differ in their attitudes to housing policy than those who have not. 

## Backgrounds

Rafael found several municipalities that publishe different types of building permits. The thing is, we need to link the building permits to residents in our survey, so it only makes sense to find building permits for municipalities that are represented in our survey. Here are the most 

```{r}
#| label: respondent-counts-by-municipality
#| cache: true
source(here("R_Scripts/1_data_import.R"))

```

Matt and Rafael and I have found several 

## Read In Each File

```{r}
#| label: import-files
# Toronto
toronto <- read.csv("Data/Building_permit_data/Toronto/Cleared Building Permits since 2017.csv")
# Ottawa
# ottawa_2020<-read_excel(path=here("Data/Building_permit_data/Ottawa/permits-2020.xls"), sheet=1)
ottawa_2021<-read_excel(path=here("Data/Building_permit_data/Ottawa/permits_2021_EN.xls"), sheet=1)
ottawa_2022<-read_excel(path=here("Data/Building_permit_data/Ottawa/Permits2022EN.xls"), sheet=1)
ottawa_2021 |> 
  bind_rows(ottawa_2022)->ottawa
# Mississauga
mississauga_sf<-read_sf("Data/Building_permit_data/Mississauga/Building_Permits.shp")
mississauga_sf |> 
  st_drop_geometry() |> 
  write_csv(here("Data/Building_permit_data/Mississauga/mississauga.csv"))
mississauga_sf |> 
  st_drop_geometry() ->mississauga

# Hamilton
hamilton_sf<-read_sf("Data/Building_permit_data/Hamilton/Building_and_Demolition_Permits_2017_to_Present.shp")
hamilton_sf |> 
  st_drop_geometry() |>
  write_csv(here("Data/Building_permit_data/Hamilton/hamilton.csv"))
hamilton_sf |> 
  st_drop_geometry() ->hamilton
# Kitchener

kitchener_sf<-read_sf(here("Data/Building_permit_data/Kitchener/Building_Permit.shp"))
kitchener_sf |> 
  st_drop_geometry() |> 
  write_csv(here("Data/Building_permit_data/Kitchener/kitchener.csv"))
kitchener_sf |> 
  st_drop_geometry() ->kitchener

#Burlington
burlington_sf<-read_sf(here("Data/Building_permit_data/Burlington/Building_Permits.shp"))
burlington_sf |> 
  st_drop_geometry() |> 
  write_csv(here("Data/Building_permit_data/Burlington/burlington.csv"))
burlington_sf |> 
  st_drop_geometry()->burlington
  
```

# Harmonization of names

The files have multiple names. This section harmonizes names for variables which are effectively the same. The first to be standardized are the dates. 

```{r}
#| label: rename-variables
library(janitor)
library(purrr)
city.list<-
  list("Toronto"=toronto, "Mississauga"=mississauga, "Burlington"=burlington, "Kitchener"=kitchener, "Hamilton"=hamilton, "Ottawa"=ottawa)
city.list |> map(clean_names)->city.list
city.list[["Mississauga"]] |> rename(completed_date=complete_d)->city.list[["Mississauga"]]
city.list[["Hamilton"]] |> rename(completed_date=completedd)->city.list[["Hamilton"]]
city.list[["Kitchener"]] |> rename(completed_date=13)->city.list[["Kitchener"]]
```

```{r}
#| label: rename-issued-date
city.list[["Mississauga"]] |> rename(issued_date=issue_date)->city.list[["Mississauga"]]
city.list[["Hamilton"]] |> rename(issued_date=issueddate)->city.list[["Hamilton"]]
city.list[["Kitchener"]] |> rename(issued_date=issue_date)->city.list[["Kitchener"]]
city.list[["Burlington"]] |> rename(issued_date=issuedate)->city.list[["Burlington"]]
```

```{r}
#| label: rename-dwelling-units-created
city.list[["Ottawa"]] |> rename(dwelling_units_created=d_u)->city.list[["Ottawa"]]
city.list[["Mississauga"]] |> rename(dwelling_units_created=res_units)->city.list[["Mississauga"]]
city.list[["Kitchener"]] |> rename(dwelling_units_created=units_crea)->city.list[["Kitchener"]]
city.list[["Burlington"]] |> rename(dwelling_units_created=resunits)->city.list[["Burlington"]]
```

```{r}
#| label: rename-permit-structure-type
city.list[["Ottawa"]] |> rename(permit_type=appl_type, structure_type=blg_type)->city.list[["Ottawa"]]
city.list[["Mississauga"]] |> rename(permit_type=scope, structure_type=bldg_type)->city.list[["Mississauga"]]
city.list[["Hamilton"]] |> rename(permit_type=workclass, structure_type=permitclas)->city.list[["Hamilton"]]
city.list[["Kitchener"]] |> rename(permit_type=permit_typ, structure_type=sub_work_t)->city.list[["Kitchener"]]
city.list[["Burlington"]] |> rename(permit_type=foldertype, structure_type=subdesc)->city.list[["Burlington"]]
```

```{r}
#| label: rename-street-address
library(stringr)
city.list[["Toronto"]] |> mutate(street_address=str_squish(paste(street_num, street_name, street_type, street_direction)))->city.list[["Toronto"]]
city.list[["Ottawa"]] |> mutate(street_address=str_squish(paste(st_number, road)))->city.list[["Ottawa"]]
city.list[["Mississauga"]] |> rename(street_address=address)->city.list[["Mississauga"]]
city.list[["Hamilton"]] |> rename(street_address=originalad)->city.list[["Hamilton"]]
city.list[["Kitchener"]] |> rename(street_address=foldername)->city.list[["Kitchener"]]
city.list[["Burlington"]] |> rename(street_address=address)->city.list[["Burlington"]]
```
# Date Filter

The first decision is the data filter. Toronto, Mississauga, Kitchener and Hamilton all of information on Completed Date.

## Residential Units Created


```{r}
#| label: units-created-histogram
#| eval: true
#| include: false


library(ggplot2)
city.list[c("Toronto", "Mississauga", "Kitchener")] |>
  map(\(x) x |>
        mutate(completed_date=as.Date(completed_date),
               dwelling_units_created=as.numeric(dwelling_units_created)) |>
        select(completed_date, dwelling_units_created, street_address, permit_type, structure_type)) |>
  bind_rows(.id="municipality") |>
  filter(completed_date>=as.Date("2021-01-01") & completed_date<as.Date("2022-05-01"))->completed

completed |>
  ggplot(aes(x=dwelling_units_created))+
  geom_histogram()+
  facet_wrap(~municipality)
```

These building permits cover almost exclusively small-scale projects that add 1 unit to properties, but there are a few tower projects in here.

I'd like to know why building permits are not issued for tower projects in general. 

```{r}
#| label: units-created-summary
completed |>
  group_by(municipality) |>
  summarize(n=n(),
            missing=sum(is.na(dwelling_units_created)),
            min=min(dwelling_units_created, na.rm=TRUE),
            q1=quantile(dwelling_units_created, 0.25, na.rm=TRUE),
            median=median(dwelling_units_created, na.rm=TRUE),
            mean=mean(dwelling_units_created, na.rm=TRUE),
            q3=quantile(dwelling_units_created, 0.75, na.rm=TRUE),
            max=max(dwelling_units_created, na.rm=TRUE),
            sd=sd(dwelling_units_created, na.rm=TRUE))
```

How many cases do we have of completed densification projects in each municipality.

```{r}
#| label: densification-count

completed |> 
  group_by(municipality) |> 
  filter(dwelling_units_created<5 |dwelling_units_created>0) |> 

  count(municipality)

```

```{r}
#| label: count-0-to-4-units
completed |>
  filter(between(dwelling_units_created, 0, 4)) |>
  group_by(municipality, street_address) |>
  summarize(permits=n(), dwelling_units_created=sum(dwelling_units_created), .groups="drop") |>
  count(municipality)
```

```{r}
#| label: units-by-address
completed |>
  group_by(municipality, street_address) |>
  summarize(permits=n(), dwelling_units_created=sum(dwelling_units_created, na.rm=TRUE), .groups="drop")->units_by_address
```

```{r}
#| label: units-per-address-variable
completed |>
  group_by(municipality, street_address) |>
  mutate(units_created_at_address=sum(dwelling_units_created, na.rm=TRUE)) |>
  ungroup()->completed
```

```{r}
#| label: inspect-addresses
set.seed(1)
units_by_address |>
  filter(permits>1 & dwelling_units_created>0) |>
  group_by(municipality) |>
  slice_sample(n=3) |>
  ungroup()->inspect_addresses
inspect_addresses |>
  knitr::kable()
completed |>
  semi_join(inspect_addresses, by=c("municipality", "street_address")) |>
  arrange(municipality, street_address, completed_date) |>
  select(municipality, street_address, completed_date, permit_type, structure_type, dwelling_units_created) |>
  knitr::kable()
```
```{r}
#| label: inspect-1-to-4-units
set.seed(1)
completed |>
  filter(units_created_at_address %in% 1:4) |>
  distinct(municipality, street_address, units_created_at_address) |>
  group_by(municipality, units_created_at_address) |>
  slice_sample(n=1) |>
  ungroup()->example_addresses
completed |>
  semi_join(example_addresses, by=c("municipality", "street_address")) |>
  arrange(municipality, units_created_at_address, street_address, completed_date) |>
  select(municipality, units_created_at_address, street_address, permit_type, structure_type, completed_date, dwelling_units_created) |>
  knitr::kable()
```


I do not think that we can *just* use the units created variable. It seems like developments that create 1 new unit include new single-familiy homes *and* additions like a new townhouse. The former is not densification, but the latter is. 

We have to poke at the variables that describe the type of development. 


```{r}
#| label: units-by-address-permit-type
completed |>
  group_by(municipality, street_address, permit_type) |>
  summarize(dwelling_units_created=sum(dwelling_units_created, na.rm=TRUE), .groups="drop")->units_by_address_permit_type
units_by_address_permit_type |> 
  filter(dwelling_units_created>0&dwelling_units_created<6) |> 
  slice_sample(n=100) |> knitr::kable()
```

```{r}
#| label: types-at-1-to-6-unit-addresses
completed |>
  filter(between(units_created_at_address, 1, 6))->mild_densification
mild_densification |>
  count(municipality, permit_type) |>
  arrange(municipality, desc(n)) |>
  knitr::kable()
mild_densification |>
  count(municipality, structure_type) |>
  arrange(municipality, desc(n)) |>
  knitr::kable()
```

```{r}
#| label: types-at-1-to-6-unit-addresses-new-units
mild_densification |>
  filter(dwelling_units_created>0)->mild_densification_new_units
mild_densification_new_units |>
  count(municipality, permit_type) |>
  arrange(municipality, desc(n)) |>
  knitr::kable()
mild_densification_new_units |>
  count(municipality, structure_type) |>
  arrange(municipality, desc(n)) |>
  knitr::kable()
```

## Densification flow chart

```{mermaid}
flowchart TD
  A["Completed permit<br/>Jan 2021 to Apr 2022"] --> B{"Units created at the address<br/>(units_created_at_address)"}
  B -- "0" --> X1["No new units:<br/>not densification"]
  B -- "more than 6" --> X2["Larger project:<br/>outside mild densification"]
  B -- "1 to 6" --> C{"Does this permit create units?<br/>(dwelling_units_created > 0,<br/>demolition permits excluded)"}
  C -- "No" --> X3["Supporting permit<br/>(plumbing, mechanical, drains, decks):<br/>describes the project, not counted"]
  C -- "Yes" --> D{"structure_type is a<br/>multi-unit or added-suite form?"}
  D -- "Yes" --> Y1["DENSIFICATION<br/>secondary suite, conversion,<br/>laneway or garden suite, multiplex"]
  D -- "No" --> E{"permit_type alters or adds to<br/>an existing building?"}
  E -- "Yes" --> Y2["DENSIFICATION<br/>unit added to an existing dwelling"]
  E -- "No: new construction" --> F{"Demolition permit<br/>at the same address?"}
  F -- "Yes" --> G{"New building has more units<br/>than one detached house?"}
  G -- "Yes" --> Y3["DENSIFICATION<br/>teardown replaced by more units"]
  G -- "No" --> X4["Replacement:<br/>one house for one house"]
  F -- "No" --> H["UNCLEAR<br/>new subdivision or infill?<br/>needs a location check"]
```

| Decision | Toronto | Mississauga | Kitchener |
|---|---|---|---|
| Multi-unit or added-suite `structure_type` | 2 Unit - Detached / Semi-detached / Townhouse; 3+ Unit - Detached / Semi-detached / Townhouse; Laneway / Rear Yard Suite; Converted House; Duplex; Triplex/Semi-Detached; Multiple Unit Building; Apartment Building | DUPLEX, 3, 4, 5 0R 6PLEX; APARTMENT (> 6 UNITS) | Duplex Dwelling (2 dwelling units); Duplex Dwelling - Condo; Semi-Duplex Dwelling; Rowhouse Duplex; Garden Suite; Apartment |
| `permit_type` alters or adds to an existing building | Small Residential Projects; Building Additions/Alterations | ALTERATION TO EXISTING BLDG; ADDITION AND ALTERATION; ADDITION TO EXISTING BLDG | none (every unit-creating permit is Residential Building (House) or (Multi)) |
| New construction `permit_type` | New Houses; New Building; Residential Building Permit | NEW BUILDING | Residential Building (House); Residential Building (Multi) |
| Demolition `permit_type` | Demolition Folder (DM) | DEMOLITION | Residential Demolition |
