---
title: "Partisan Consensus and its Limits in Ontario Housing Policy"
author: 
  - "Simon Kiss"
  - "Anthony Piscitelli"
  - "Matthew Arp"
date: "2023-10-16"
output: 
  powerpoint_presentation: 
    reference_doc: ~/lispop_template.pptx
bibliography: housing.bib

---



## Ontario Provincial Election Survey

```{r setup, include=FALSE, warning=F, message=F, cache=FALSE, include=F}
knitr::opts_chunk$set(echo = FALSE, cache=F, warning=F, message=F, dev="png")
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r load-data, results="hide", include=F}
source("R_Scripts/4_pca.R")

```

```{r set-theme, echo=F}
my_theme <- function(base_size = 14) {
  theme_bw(base_size = base_size) %+replace%
    theme(
      legend.position="bottom",
      plot.title=element_text(size=18)
    )
}

theme_set(my_theme(base_size = 20))
```

- LISPOP's mandate to support research, training and knowledge mobilization of public opinion and policy
- Survey of 2000 voters during 2022 Ontario Provincial Election
- Field dates between May 18th and May 30th, 2022
- Standard political variables (e.g. federal and provincial vote, federal and provincial partisanship) as well as heavy focus on housing-specific variables
- Data publicly available [here](https://doi.org/10.5683/SP3/FFTYDV) for sharing

::: notes
- Support for LISPOP comes from Office of Research Services
- Strength of Laurier in all aspects of public opinion training and research, e.g. survey design, analysis
- regularly commission surveys on important issues, we conducted an OPES 2018 and wanted to repeat that in 2022
- clear Housing was going to be an issue, focus of this survey
- emphasize dataset publicly available
:::

## Context of the 2022 Election Campaign

:::::::::::::: {.columns}
::: {.column width="60%"}

- Provincial government commissioned an expert panel to make recommendations on housing affordability
- Signature recommendation to build 1.5 million homes
- subsequently included in the NDP and Liberal platforms

:::
::: {.column width="40%"}
```{r mip, ft.align="center", ft.top=2,ft.left=7}
on22 %>% 
  count(Topic) %>% 
  mutate(Percent=n/sum(n)*100) %>% 
  arrange(desc(Percent)) %>% 
  filter(Percent>4) %>% 
  flextable() %>% 
  bg(., i=3, bg="yellow") %>% 
  colformat_double(digits = 0) %>% 
  width(j=1, width=4)

```
:::
::::::::::::::


::: notes
- Prior to the campaign, expert panel announced to make recommendations
- Signature policy was to create 1.5 million homes over 10 years
- Cross-partisan consensus on the need to increase supply
  - Implementation is harder than you might think
  - The data here will show how this support seeming consensus is fragile
  
- Governments can't just wave a wand and "increase supply". Requires financing, cooperation from developers, municipalities, rule changes, each of which can present place obstacles to emerge
:::


## Solutions


```{r means-graph, fig.width=11, fig.height=6}
on22 %>% 
  select(Q33a_1_x:Q80_6_x) %>% 
  pivot_longer(cols=everything(),names_to="variable") %>% 
  left_join(solution_var_labels, by="variable") %>% 
  group_by(label) %>% 
  summarize(Average=mean(value, na.rm=T), sd=sd(value, na.rm=T), n=n(),se=sd/sqrt(n)) %>% 
  ggplot(., aes(x=Average, y=fct_reorder(label, Average)))+
  geom_pointrange(size=1.2,aes(xmin=Average-(1.96*se), xmax=Average+(1.96*se)))+
  labs(x="Score (0=Strongly Oppose\n1=Strongly Support)", 
       title=str_wrap("Solutions To Address Housing Price Increase",60), y="")+
  xlim(c(0,1))+
 # scale_y_discrete(labels=function(x) str_wrap(x, 20))+
  geom_vline(xintercept=0.5, linetype=2)
```


## Ideology in housing policy

:::::::::::::: {.columns}
::: {.column width="40%"}
- Two clusters of policies: 

  1. Government regulation
  2. Reducing barriers to increasing supply (Market forces)

- These cluster mirror two fundamental ways of accomplishing collective goals: markets and states (or bureaucracy) [@samuelsonPureTheoryPublic1954]
- In this case, government regulation is noticeably more popular 
- The general goal of increasing supply is actually correlated with *both* clusters
:::
::: {.column width="60%"}
```{r pca-results, ft.left=7} 

library(flextable)
pca_loadings<-data.frame(fa.sort(support_pca2$loadings)[1:12,])
pca_loadings %>% 
  add_rownames() %>% 
  flextable() %>% 
  colformat_double(., digits=2) %>% bold(., ~RC1 >0.4, j="RC1") %>% 
  bold(., ~RC2 >0.4, j="RC2")   %>% 
  width(., j="rowname", width=3) %>% 
    width(., j="RC1", width=1) %>% 
  width(., j="RC2", width=1) %>% 
  set_header_labels(values=list(rowname="Question", RC1="Government", RC2="Market")) 

```
:::
::::::::::::::

## Ideology in housing policy

:::::::::::::: {.columns}
::: {.column width="40%"}
- Two clusters of policies: 

  1. Government regulation
  2. Reducing barriers to increasing supply (Market forces)

- These cluster mirror two fundamental ways of accomplishing collective goals: markets and states (or bureaucracy) [@samuelsonPureTheoryPublic1954]
- In this case, government regulation is noticeably more popular 
- The general goal of increasing supply is actually correlated with *both* cluster
:::
::: {.column width="60%"}
```{r pca-results-1, ft.left=7} 

library(flextable)
pca_loadings<-data.frame(fa.sort(support_pca2$loadings)[1:12,])
pca_loadings %>% 
  add_rownames() %>% 
  flextable() %>% 
  colformat_double(., digits=2) %>% bold(., ~RC1 >0.4, j="RC1") %>% 
  bold(., ~RC2 >0.4, j="RC2")   %>% 
 bg(., i=11, bg="yellow") %>% 
  width(., j="rowname", width=3) %>% 
    width(., j="RC1", width=1) %>% 
  width(., j="RC2", width=1) %>% 
  set_header_labels(values=list(rowname="Question", RC1="Regulation", RC2="Market")) 

```
:::
::::::::::::::


## Importance of partisan cues

:::::::::::::: {.columns}
::: {.column width="20%"}

- Voters often take cues from parties they identify with as to what positions they should hold [@zallerNatureOriginsMass1992]
- Cross-partisan consensus in the platforms is evident in 
:::

::: {.column width="80%"}
```{r solutions-partisanship, fig.width=10, fig.height=8}
on22 %>% 
  #Pick the variables working with
  select(Q33a_1_x:Q80_6_x, partisanship) %>% 
  #pivot them longer, except for the Sample variable
  pivot_longer(., cols=-partisanship, names_to=c("variable")) %>% 
  group_by(partisanship, variable) %>% 
  filter(!is.na(value)) %>% 
  filter(partisanship!="Green") %>% 
  # filter(partisanship!=5) %>% #Filter out R's that identify as "Other" 
  # filter(partisanship!=6) %>%# Filter out R's that identify as "None of these"
  summarize(average=mean(value), sd=sd(value), n=n(), se=sd/sqrt(n)) %>% 
  left_join(., solution_var_labels)  %>% 
  filter(partisanship!="Independent") %>% 
  ggplot(., aes(y=fct_reorder(label, average), x=average, col=partisanship))+
  xlim(c(0,1))+geom_point()+
 geom_pointrange(size=1,aes(xmin=average-(1.96*se),
                            xmax=average+(1.96*se)))+
 scale_color_manual(name=c('Partisanship'), values=c("blue", "orange", "darkred"))+
  labs(x=str_wrap("Score (0=Strongly Oppose, 1=Strongly Support)", 30), 
       y="")+
  geom_vline(xintercept=0.5, linetype=2)+
  guides(col=guide_legend(ncol=3))
```
:::
::::::::::::::


## Popularity of the clusters

::: notes
Overall, if we compare the clusters by partisanship, it's clear that a lot of the difference is driven by left-wing support for government regulation. PC voters are less enthusiastic than the other parties to do support either cluster.
:::

```{r averages, fig.width=10, fig.height=8}
on22 %>% 
  select(partisanship, contains("_Average")) %>% 
  pivot_longer(-partisanship) %>% 
  group_by(partisanship, name) %>% 
  summarize(Average=mean(value)) %>% 
  rename(Partisanship=partisanship, Measure=name) %>% 
  filter(Partisanship!="Green") %>% 
  mutate(Measure=fct_recode(Measure, Government="Government_Average", Market="Market_Average")) %>% 
  ggplot(., aes(x=Average, y=Measure, col=Partisanship))+
  geom_point(size=5,position=position_dodge(width=0.2))+xlim(c(0,1))+
  guides(col=guide_legend(ncol=2))+scale_color_manual(values=c("darkblue", "orange", "darkred", "grey"))+
  geom_vline(linetype=2, xintercept=0.5)


```

## Conclusion

1. Cross-partisan consensus on the need to increase housing supply
2. Public opinion remains protective of single-family homes
3. Combined with public opposition to sprawl leaves the impression of remarkable challenges to adding to supply, despite #1

::: notes
1. Cross-partisan consensus on the need to increase housing supply
  - This only goes so far
2. Public opinion remains protective of single-family homes
    - Government action on prices remains more popular than government action which might increase housing supply without contributing to urban sprawl
    
    3. Combined with public opposition to sprawl (see blowback on Greenbelt) leaves the impression of remarkable challenges to adding to supply, despite #1
    
:::

## References

